Multi-wavelength Radio Continuum Emission Studies of Dust-free Red Giants

### 2013-07-11 aj_submission_2/ms.tex
Jul 11th: Almost ready to send off

In order to understand the mechanisms which drive the $10^{-9}-10^{-11}$ $M_{\odot}$ yr$^{-1}$ mass loss rates from evolved spectral-type K through mid-M stars, an understanding of the dynamics and thermodynamics of their atmospheres is essential. An important discovery in late-type evolved stellar atmospheres resulted from the first ultraviolet survey of such stars using the International Ultraviolet Explorer (IUE). The survey revealed a `transition region dividing line' in the giant branch near spectral type K1 III which separates these stars based on the properties of their atmospheres \citep{1979ApJ...229L..27L}. Stars blueward of the dividing line were found to possess chromospheres and transition regions like the Sun, while stars on the red side were found to possess chromospheres and cool winds. X-ray observations showed that this dividing line extended to coronal emission \citep{1981ApJ...250..293A}. Around the same time, another class of late-type evolved star emerged which showed signs of possessing both a transition region and a cool wind \citep[e.g.,][]{1982A&A...107..292R}. Many of these so called `hybrid atmosphere' stars now also show evidence for coronal emission, albeit much weaker than on the blue side of the dividing line \citep{1997ApJ...491..876A,2005ApJ...622..629D}. Understanding the nature of the atmospheric structure of late-type evolved stars will ultimately lead to a broader understanding of the mass-loss process.

@@

Although studies of wind-scattered UV and optical line profiles have provided clues to the mass loss rates and radial distribution of the mean and turbulent velocity fields, the thermal structure remains poorly constrained. In the UV, the source function, $S_{\nu}$, of electron collisionally excited emission lines is sensitive to electron temperature, $T_{\rm{e}}$, (i.e., $S_{\nu} \propto e^{-h\nu /kT_{\rm{e}}}/\sqrt{T_{\rm{e}}}$). Therefore, a localized hot plasma component in a dynamic atmosphere can completely dominate the temporally and spatially averaged emission and hence not reflect the mean radial electron temperature distribution. At radio wavelengths however, the source function is thermal and is just the Rayleigh-Jeans tail of the Planck function, which is linear in electron temperature (i.e., $S_{\nu} = {2kT_{\rm{e}}\nu ^2 /c^2}$). This should give a more appropriate estimate of the mean radial electron temperature. It is this value that controls the atomic level populations and ionization of the mean plasma which is needed to quantify the implied thermal heating supplied to the  wind by the unknown driving source/sources, allowing constraints on potential mass loss mechanisms to be derived.

In the cm-radio regime the radio opacity, $ \kappa _{\lambda}$, strongly increases with wavelength (i.e., $ \kappa _{\lambda} \propto \lambda ^{2.1}$) and so the longer wavelengths sample the extended layers of a stars atmosphere, thus	 providing us with spatial information about the star's mass outflow region. The NRAO\footnote{The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc.} Karl G. Jansky Very Large Array (VLA) is sensitive to over three orders of magnitude in continuum optical depth, $\tau _{\lambda}$, ($\tau _{20 \ \rm{cm}}/\tau _{0.7 \ \rm{cm}} \approx 10^3$) and provides an area-averaged sweep through the wind acceleration zone of evolved late-type stars. The thermodynamic properties in this spatial region control the ionization in the far wind because the ionization balance, which also controls the cooling rates, becomes \textit{frozen-in} at large radii due to advection. Furthermore, it is these outer extended regions of the star's atmosphere that contribute to the commonly seen P Cygni line profiles in the UV. In these profiles the line-of-sight absorption caused by the star's wind is superimposed on the blueshifted scattered emission. Thus, centimeter radio continuum observations can provide a test of models based on these UV profiles. In this paper we directly compare our new VLA observations with atmospheric models derived from UV analysis.

@@

In order to avoid the assumed additional complexities of a companion, we have selected two single luminosity class III red giants: Arcturus ($\alpha$ Boo: K2 III) and Aldebaran ($\alpha$ Tau: K5 III). These nearby red giants have been extensively studied at other wavelengths and their stellar parameters, which are briefly summarized in Table 1, are accurately known. Both of these late-type giants have `hybrid atmospheres' as they show evidence for both coronal/transition region activity and strong winds. Even though they are slow rotators, a mean longitudinal magnetic field (albeit weak: $B = 0.65 \pm 0.26$ G) has been detected on $\alpha$ Boo \citep{2011A&A...529A.100S} along with a possible magnetic cycle with a period of $\geq$14 yr \citep{2008ApJ...679.1531B}. Also, the detection of O $\textsc{vi}$ in $\alpha$ Tau \citep{2005ApJ...622..629D} indicates magnetic activity in its atmosphere. These stars are predicted to be point sources at all frequencies between 1 and 50 GHz in all VLA configurations so our radio observations measure their total flux density, $F_{\nu}$. Moreover, both stars have existing semi-empirical 1-D chromospheric and wind models which we directly compare to our data in this paper.

@@

\subsection{Existing Atmospheric Models} \label{disc2}
One of the most important diagnostic features indicating mass outflows in late-type evolved stars are the blue shifted absorption components present in the Ca\,\textsc{ii} H and K and Mg\,\textsc{ii} h and k resonance lines. Figure \ref{fig:fig0} shows one of the two chromosphere and wind models of $\alpha$ Boo \cite[`model A']{1985pssl.proc..351D} which is based on the Mg\,\textsc{ii} k 2796 $\rm{\AA}$ emission line observed with the International Ultraviolet Explorer. The line was modeled by solving the radiative transfer equation in a spherical co-moving frame and the effects of partial redistribution \citep[e.g.,][]{1983ApJ...273..299D} were taken into account. Both of Drake's atmospheric models contain the photospheric model of \cite{1975ApJ...200..660A}, predict the wind to reach a terminal velocity of 35 - 40 km s${}^{-1}$ by 2 $R _{\star}$, and reach a maximum microturbulence of 5 km s$^{-1}$. They contain a broad temperature plateau with $T_{\rm{e}}$ $\approx$ 8,000 K between 1.2 and $\sim$20 $R _{\star}$ with a cooler region further out, and hydrogen is 50\% ionized. We compute the radio spectrum from these models assuming spherical 1-D geometry \citep{1994MNRAS.268..894H} with the free-free Gaunt factors from \cite{1988ApJ...327..477H}. The radiative transfer equation is solved using the Feautrier technique \citep{1978stat.book.....M} and the boundary condition is determined by ensuring the atmosphere is optically thick at the deepest layers. \cite{1985pssl.proc..351D} predicts that their atmospheric model would produce a flux density value of 0.4 mJy at 6 cm and encouragingly, our radio spectrum reproduces this value. Departures from spherical symmetry are to be expected in magnetic stellar atmospheres. For example, $\alpha$ Boo has an inclination axis of 58$^{\circ} \pm$25$^{\circ}$ \citep{2006PASP..118.1112G} and a global magnetic dipole could cause density variations between the equator and the polar regions. Despite this fact, the study of a  spherically symmetric atmosphere forms the basis of understanding the more complex environments in real stellar atmospheres.

@@

Figure \ref{fig:fig1} shows the resulting predicted radio spectrum between 1 GHz and 1 THz for $\alpha$ Boo from these chromosphere and wind models (green line). At high frequencies the radio spectra produced by these models have a blackbody-like slope (i.e., $\sim \nu ^{2}$) as a result of the small ion density scale heights close to the star where the temperature is changing slowly. At low frequencies however, where the Drake models predict the wind to have constant velocity, ionization fraction, and temperature, the slopes approaches the well known $\sim\nu ^{0.6}$ limit \citep{1975MNRAS.170...41W,1975AA....39..217O,1975AA....39....1P}. The paucity and, in some cases low S/N, of previous observations made it difficult to discern the validity of this model prior to our multi-frequency study of $\alpha$ Boo. Our new data reveal significant deviations from the semi-empirical model at both low and high frequencies (in this case below $\sim$8 GHz and above $\sim$25 GHz). At high frequencies our VLA data indicates a flux excess which is in agreement with previous mm-observations. The discrepancy at low frequencies may be due to a lower ionization fraction in the wind, or a lower mass loss rate than that used in the Drake model.

In Figure \ref{fig:fig2} we plot the expected radio spectrum of $\alpha$ Tau based on the semi-empirical 1-D chromosphere and transition region model of \cite{1999MNRAS.302...37M} embedded in the 1-D wind model of \cite{1998ApJ...503..396R}. The semi-empirical McMurry model was created by using the radiative transfer code MULTI \citep{carlsson_1986} to reproduce the fluxes of collisionally excited C\,\textsc{i}, C\,\textsc{ii}, Si\,\textsc{iii}, Mg\,\textsc{ii}, and C\,\textsc{iv} lines in a plane-parallel, hydrostatic, one-component atmosphere. It contains the photospheric model of \cite{1973ApJ...180...81J} and reaches a maximum temperature of 10$^{5}$ K at 1.2 $R_{\star}$. As it does not contain a wind outflow, we use Robinson et al.'s wind characteristics beyond 1.2 $R_{\star}$ to describe the outflow velocity. In this wind model, the wind reaches $\sim$80\% of its terminal value of 30 km s$^{-1}$ by 3 $R_{\star}$. The Robinson et al. wind characteristics are based on matching the Fe\,\textsc{ii} 2755 $\rm{\AA}$ line and the O\,\textsc{i} triplet near 1304 $\rm{\AA}$ with a simplified wind model using the SEI computer code \citep{1987ApJ...314..726L}. We assume the wind to have a constant temperature of 10,000 K and have a constant ionization fraction, $x_{e}$, of 0.5 throughout, based on the ionization fraction at the corresponding temperature in the McMurry model.

The radio flux densities at high frequencies (i.e., $\nu >$ 30 GHz) are overestimated by the combination of both atmospheric models, although this approach does well in reproducing the VLA flux densities below 30 GHz. The VLA, Institut de Radioastronomie Millim\'{e}trique (IRAM) 30 m-telescope and Berkeley Illinois Maryland Association (BIMA) continuum flux densities confirm that this model predicts a flux excess at even higher frequencies. One possible explanation for this is that the inner atmosphere contains extensive amounts of cooler gas than that predicted by the 1-D static chromospheric model of McMurry. This scenario agrees with the findings of \cite{1994ApJ...423..806W} who conclude that cool regions exist close to the stellar surface with large ($> 99 \%$) filling factors i.e., a thermally bifurcated CO-mosphere \citep{1996IAUS..176..371A}. The wind which we have overlain on top of the McMurry chromosphere and transition region is found to be optically thin at nearly all VLA wavelengths, and only contributes a very small flux density at the longest wavelengths. As our model matches the data reasonably well below 30 GHz we conclude that $\alpha$ Tau's wind is optically thin and the VLA radio emission at all wavelengths emanates from the inner atmosphere.

@@

Recently, \cite{2013A&A...553A...3O} has detected a layer of CO in the outer atmosphere of $\alpha$ Tau (i.e. a so-called MOLsphere) which extends out to $\sim$2.5 $R_{\star}$, has a temperature of $1500 \ \pm 200$ K, and a CO column density of $\sim$1$\times 10^{20}$ cm$^{-2}$. They were unable to constrain the geometrical thickness $\Delta L$, of the MOLsphere from the data however, and arbitrarily set it to 0.1 $R_{\star}$. It can be shown that such a MOLsphere would have an optical depth of $\tau _{\rm{6\,cm}} = 4.6$ at C-band and would produce a corresponding flux density of  0.06 mJy which is considerably lower than our high S/N measurement of 0.15 mJy. Here, we have conservatively assumed that the electrons in this region of the atmosphere come from singly ionized metals and have an abundance of $\sim$1$\times 10^{-5} n_{\rm{H}}$, where $n_{\rm{H}}$ is the total hydrogen number density. We have also assumed the CO filling factor to be unity. The disagreement in values between our radio data and the predicted flux for an optically thick disk could mean that the MOLsphere is optically thin at long VLA wavelengths. It can be shown that the MOLsphere becomes optically thin at C-band (i.e. $\tau _{\rm{6\,cm}} < 1$) for $\Delta L >$0.45 $R_{\star}$, so if the long VLA wavelength radio emission from $\alpha$ Tau comes from a region much closer in to the star then the MOLsphere either has a geometrical width  $>$0.45 $R _{\star}$ or has a filling factor less than unity.

@@

Equation 1 can be used in conjunction with our new spectral index for each star to calculate the density and temperature coefficients that may describe their outflows. The combinations of the electron temperature and density coefficients are shown for each star in Figure 5 ($\alpha$ Boo is represented by the solid line and $\alpha$ Tau by the dash-dotted line) along with the coefficients obtained by assuming either an isothermal flow ($n=0$) or a constant velocity flow ($p=2$). One potential explanation for spectral indices of stellar outflows being larger than 0.6 is that the wind is still accelerating in the radio emitting region, if the thermal gradients are assumed to be small. If we ignore thermal gradients completely, then the density coefficients are $p  =2.7$ and 5.5 for $\alpha$ Boo and $\alpha$ Tau, respectively. From mass conservation arguments assuming a steady flow, the power law coefficients for the velocity profiles of each star can be found, ie., $v(r) \propto r^{p-2}$. For $\alpha$ Boo we find $v(r) \propto r^{0.7}$ while for $\alpha$ Tau we find $v(r) \propto r^{3.5}$. This suggests that our long VLA wavelengths may probe a steep acceleration region for $\alpha$ Tau's outflow but for $\alpha$ Boo, may probe a region where its wind is close to its terminal velocity.

@@

The assumption of completely ignoring thermal gradients in a stellar outflow is probably unreliable however. It is likely that some form of Alfv\`{e}n waves are required to lift the material out of the gravitational potential as suggested by \cite{1980ApJ...242..260H}. These waves need damping lengths which are much larger than the density scale height $H$ (where $H \sim 0.005 R_{\star}$ for our targets), in order to lift the material out of the gravitational potential but $\lesssim 1 \, R_{\star}$ in order to avoid wind terminal velocities greater than those observed \citep[e.g.][]{1983ApJ...275..808H}. It has also been shown that these waves are expected to produce substantial heating deposition near the base of the wind \citep{1982ApJ...261..279H}. If the long wavelength radio emission from $\alpha$ Tau is indeed emanating from the wind acceleration region, then the dissipation of these Alfv\`{e}n waves may introduce thermal gradients in this region and its velocity profile will not be described by $v(r) \propto r^{3.5}$. Furthermore, diverging flow geometries have been invoked as a more realistic representation of stellar atmospheres \citep{1982ApJ...257..264H,1989A&A...209..327J,2006SSRv..122..181V} and so one could write the continuity equation for a steady flow as 
\begin{equation}
\dot{M} _{\star} = f(r)4\pi r^2 \rho (r) v(r),
\end{equation}
where $f(r)$ is a function describing the divergence from a purely radial flow [i.e., $f(r)=1$], $\dot{M} _{\star}$ is the stellar mass loss rate, and $\rho(r)$ is the gas density at $r$. If the non-radial expansion term can be described by a power-law $f(r) \propto r^s$, where $s>1$ indicates super-radial expansion, then $v(r) \propto r^{3.5-s}$ would describe the velocity profile of $\alpha$ Tau (ignoring thermal gradients and assuming a constant ionization fraction). Further out in the wind, where it has reached its terminal velocity, one would also expect a thermal gradient (but now of opposite sign) due to adiabatic expansion and line cooling. If the long wavelength radio emission emanates from this region of the wind then Equation \ref{eq:eq1} provides us with a direct estimate of the temperature coefficient as we can assume the density coefficient is $p=2$. This may be the case for our long VLA wavelength measurements of $\alpha$ Boo in which case $T_{e}(r) \propto r^{-1.65}$.

@@

1. A constant velocity mass outflow, i.e., $n_{\rm{H}}(r)=C/r^2$ where $n_{\rm{H}}$ is the total hydrogen number density and $C$ is a constant proportional to the ratio of the mass loss rate divided by the terminal velocity. For $\alpha$ Boo, $C = 1.7 \times 10^{32} $ cm$^{-1}$ assuming a wind velocity of 35 km s$^{-1}$.

>>>>

In order to understand the mechanisms which drive the $10^{-9}-10^{-11}$ $M_{\odot}$ yr$^{-1}$ mass loss rates from evolved spectral-type K through mid-M stars, an understanding of the dynamics and thermodynamics of their atmospheres is essential. An important discovery in late-type evolved stellar atmospheres resulted from the first ultraviolet survey of such stars using the International Ultraviolet Explorer (IUE). The survey revealed a `transition region dividing line' in the giant branch near spectral type K1 III which separates these stars based on the properties of their atmospheres \citep{1979ApJ...229L..27L}. Stars blueward of the dividing line were found to possess chromospheres and transition regions like the Sun, while stars on the red side were found to possess chromospheres and cool winds. X-ray observations showed that this dividing line extended to coronal emission \citep{1981ApJ...250..293A}. Around the same time, another class of late-type evolved star emerged which showed signs of possessing both a transition region and a cool wind \citep[e.g.,][]{1982A&A...107..292R}. Many of these so-called `hybrid atmosphere' stars now also show evidence for coronal emission, albeit much weaker than on the blue side of the dividing line \citep{1997ApJ...491..876A,2005ApJ...622..629D}. Understanding the nature of the atmospheric structure of late-type evolved stars will ultimately lead to a broader understanding of the mass-loss process.

@@

Although studies of wind-scattered UV and optical line profiles have provided clues to the mass loss rates and radial distribution of the mean and turbulent velocity fields, the thermal structure remains poorly constrained. In the UV, the source function $S_{\nu}$, of electron collisionally excited emission lines is sensitive to electron temperature $T_{\rm{e}}$, (i.e., $S_{\nu} \propto e^{-h\nu /kT_{\rm{e}}}/\sqrt{T_{\rm{e}}}$). Therefore, a localized hot plasma component in a dynamic atmosphere can completely dominate the temporally and spatially averaged emission and hence not reflect the mean radial electron temperature distribution. At radio wavelengths however, the source function is thermal and is just the Rayleigh-Jeans tail of the Planck function, which is linear in electron temperature (i.e., $S_{\nu} = {2kT_{\rm{e}}\nu ^2 /c^2}$). This should give a more appropriate estimate of the mean radial electron temperature. It is this value that controls the atomic level populations and ionization of the mean plasma which is needed to quantify the implied thermal heating supplied to the  wind by the unknown driving source/sources, allowing constraints on potential mass loss mechanisms to be derived.

In the cm-radio regime the radio opacity $ \kappa _{\lambda}$, strongly increases with wavelength (i.e., $ \kappa _{\lambda} \propto \lambda ^{2.1}$) and so the longer wavelengths sample the extended layers of a stars atmosphere, thus	 providing us with spatial information about the star's mass outflow region. The NRAO\footnote{The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc.} Karl G. Jansky Very Large Array (VLA) is sensitive to over three orders of magnitude in continuum optical depth, $\tau _{\lambda}$, ($\tau _{20 \ \rm{cm}}/\tau _{0.7 \ \rm{cm}} \approx 10^3$) and provides an area-averaged sweep through the wind acceleration zone of evolved late-type stars. The thermodynamic properties in this spatial region control the ionization in the far wind because the ionization balance, which also controls the cooling rates, becomes \textit{frozen-in} at large radii due to advection. Furthermore, it is these outer extended regions of the star's atmosphere that contribute to the commonly seen P Cygni line profiles in the UV. In these profiles the line-of-sight absorption caused by the star's wind is superimposed on the blueshifted scattered emission. Thus, centimeter radio continuum observations can provide a test of models based on these UV profiles. In this paper we directly compare our new VLA observations with atmospheric models derived from UV analysis.

@@

In order to avoid the assumed additional complexities of a companion, we have selected two single luminosity class III red giants: Arcturus ($\alpha$ Boo: K2 III) and Aldebaran ($\alpha$ Tau: K5 III). These nearby red giants have been extensively studied at other wavelengths and their stellar parameters, which are briefly summarized in Table 1, are accurately known. Both of these late-type giants have `hybrid atmospheres' as they show evidence for both coronal/transition region activity and strong winds. Even though they are slow rotators, a mean longitudinal magnetic field (albeit weak: $B = 0.65 \pm 0.26$ G) has possibly been detected for $\alpha$ Boo \citep{2011A&A...529A.100S} along with a possible magnetic cycle with a period of $\geq$14 yr \citep{2008ApJ...679.1531B}. Also, the detection of O $\textsc{vi}$ in $\alpha$ Tau \citep{2005ApJ...622..629D} indicates magnetic activity in its atmosphere. These stars are predicted to be point sources at all frequencies between 1 and 50 GHz in all VLA configurations so our radio observations measure their total flux density, $F_{\nu}$. Moreover, both stars have existing semi-empirical 1-D chromospheric and wind models which we directly compare to our data in this paper.

@@

\subsection{Existing Atmospheric Models} \label{disc2}
One of the most important diagnostic features indicating mass outflows in late-type evolved stars are the blue shifted absorption components present in the Ca\,\textsc{ii} H and K and Mg\,\textsc{ii} h and k resonance lines. Figure \ref{fig:fig0} shows one of the two chromosphere and wind models of $\alpha$ Boo \cite[`model A']{1985pssl.proc..351D} which is based on the Mg\,\textsc{ii} k 2796 $\rm{\AA}$ emission line observed with the IUE telescope. The line was modeled by solving the radiative transfer equation in a spherical co-moving frame and the effects of partial redistribution \citep[e.g.,][]{1983ApJ...273..299D} were taken into account. Both of Drake's atmospheric models contain the photospheric model of \cite{1975ApJ...200..660A}, predict the wind to reach a terminal velocity of 35 - 40 km s${}^{-1}$ by 2 $R _{\star}$, and reach a maximum microturbulence of 5 km s$^{-1}$. They contain a broad temperature plateau with $T_{\rm{e}}$ $\approx$ 8,000 K between 1.2 and $\sim$20 $R _{\star}$ with a cooler region further out, and hydrogen is $\sim50$\% ionized. We compute the radio spectrum from these models assuming spherical 1-D geometry \citep{1994MNRAS.268..894H} with the free-free Gaunt factors from \cite{1988ApJ...327..477H}. The radiative transfer equation is solved using the Feautrier technique \citep{1978stat.book.....M} and the boundary condition is determined by ensuring the atmosphere is optically thick at the deepest layers. \cite{1985pssl.proc..351D} predicts that their atmospheric model would produce a flux density value of 0.4 mJy at 6 cm and encouragingly, our radio spectrum reproduces this value. Departures from spherical symmetry are to be expected in magnetic stellar atmospheres. For example, $\alpha$ Boo has an inclination axis of 58$^{\circ} \pm$25$^{\circ}$ \citep{2006PASP..118.1112G} and a global magnetic dipole could cause density variations between the equator and the polar regions. Despite this fact, the study of a  spherically symmetric atmosphere forms the basis of understanding the more complex environments in real stellar atmospheres.

@@

Figure \ref{fig:fig1} shows the resulting predicted radio spectrum between 1 GHz and 1 THz for $\alpha$ Boo from these chromosphere and wind models (green line). At high frequencies the radio spectra produced by these models have a blackbody-like slope (i.e., $\sim \nu ^{2}$) as a result of the small ion density scale heights close to the star where the temperature is changing slowly. At low frequencies however, where the Drake models predict the wind to have constant velocity, ionization fraction, and temperature, the slopes approaches the well known $\sim\nu ^{0.6}$ limit \citep{1975MNRAS.170...41W,1975AA....39..217O,1975AA....39....1P}. The paucity and, in some cases low S/N, of previous observations made it difficult to discern the validity of this model prior to our multi-frequency study of $\alpha$ Boo. Our new data reveal significant deviations from the semi-empirical model at both low and high frequencies (in this case below $\sim$8 GHz and above $\sim$25 GHz). At high frequencies our VLA data indicates a flux excess which is in agreement with previous mm-observations. This may be due to larger chromospheric ion densities or to the possible presence of transition region plasma not accounted for in the Drake model. The discrepancy at low frequencies may be due to a lower ionization fraction in the wind, or a lower mass loss rate than that used in the Drake model.

In Figure \ref{fig:fig2} we plot the expected radio spectrum of $\alpha$ Tau based on the semi-empirical 1-D chromosphere and transition region model of \cite{1999MNRAS.302...37M} embedded in the 1-D wind model of \cite{1998ApJ...503..396R}. The semi-empirical McMurry model was created by using the radiative transfer code MULTI \citep{carlsson_1986} to reproduce the fluxes of collisionally excited C\,\textsc{i}, C\,\textsc{ii}, Si\,\textsc{iii}, Mg\,\textsc{ii}, and C\,\textsc{iv} lines in a plane-parallel, hydrostatic, one-component atmosphere. It contains the photospheric model of \cite{1973ApJ...180...81J} and reaches a maximum temperature of 10$^{5}$ K at 1.2 $R_{\star}$. As it does not contain a wind outflow, we use Robinson et al.'s wind characteristics beyond 1.2 $R_{\star}$ to describe the outflow velocity. In this wind model, the wind reaches $\sim$80\% of its terminal value of 30 km s$^{-1}$ by 3 $R_{\star}$. The Robinson et al. wind characteristics are based on matching the Fe\,\textsc{ii} 2755 $\rm{\AA}$ line and the O\,\textsc{i} triplet near 1304 $\rm{\AA}$ with a simplified wind model using the SEI computer code \citep{1987ApJ...314..726L}. We assume the wind to have a constant temperature of 10,000 K and have a constant ionization fraction of 0.6 throughout, based on the ionization fraction at the corresponding temperature in the McMurry model.

The radio flux densities at high frequencies (i.e., $\nu >$ 30 GHz) are overestimated by the combination of both atmospheric models, although this approach does well in reproducing the VLA flux densities below 30 GHz. The VLA, Institut de Radioastronomie Millim\'{e}trique (IRAM) 30 m-telescope and Berkeley Illinois Maryland Association (BIMA) continuum flux densities confirm that this model predicts a flux excess at even higher frequencies. One possible explanation for this is that the inner atmosphere contains extensive amounts of cooler gas than that predicted by the 1-D static chromosphere and transition region model of McMurry. This scenario agrees with the findings of \cite{1994ApJ...423..806W} who conclude that cool regions exist close to the stellar surface with large ($> 99 \%$) filling factors i.e., a thermally bifurcated CO-mosphere \citep{1996IAUS..176..371A}. The wind which we have overlain on top of the McMurry chromosphere and transition region is found to be optically thin at nearly all VLA wavelengths, and only contributes a very small flux at the longest wavelengths. As our model matches the data reasonably well below 30 GHz we conclude that $\alpha$ Tau's wind is optically thin and the VLA radio emission at all wavelengths emanates from the inner atmosphere.

@@

Recently, \cite{2013A&A...553A...3O} has detected a layer of CO in the outer atmosphere of $\alpha$ Tau (i.e. a so-called MOLsphere) which extends out to $\sim$2.5 $R_{\star}$, has a temperature of $1500 \ \pm 200$ K, and a CO column density of $\sim$1$\times 10^{20}$ cm$^{-2}$. They were unable to constrain the geometrical thickness $\Delta L$, of the MOLsphere from the data however, and arbitrarily set it to 0.1 $R_{\star}$. It can be shown that such a MOLsphere would have an optical depth of $\tau _{\rm{6\,cm}} = 4.6$ at C-band and would produce a corresponding flux density of  0.06 mJy which is considerably lower than our high S/N measurement of 0.15 mJy. Here, we have conservatively assumed that the electrons in this region of the atmosphere come from singly ionized metals and have an abundance of $\sim$1$\times 10^{-5} n_{\rm{H}}$, where $n_{\rm{H}}$ is the total hydrogen number density. We have also assumed the CO filling factor to be unity. The disagreement in values between our radio data and the predicted flux for an optically thick disk could mean that the MOLsphere is optically thin at long VLA wavelengths and that the radio emission emanates from the more ionized material closer to the star. It can be shown that the MOLsphere becomes optically thin at C-band (i.e. $\tau _{\rm{6\,cm}} < 1$) for $\Delta L >$0.45 $R_{\star}$, so if the long VLA wavelength radio emission from $\alpha$ Tau comes from a region closer in to the star, then the MOLsphere either has a geometrical width  $>$0.45 $R _{\star}$ or has a filling factor less than unity.

@@

Equation 1 can be used in conjunction with our new spectral index for each star to calculate the density and temperature coefficients that may describe their outflows. The combinations of the electron temperature and density coefficients are shown for each star in Figure 5 ($\alpha$ Boo is represented by the solid line and $\alpha$ Tau by the dash-dotted line) along with the coefficients obtained by assuming either an isothermal flow ($n=0$) or a constant velocity flow ($p=2$). One potential explanation for spectral indices of stellar outflows being larger than 0.6 is that the wind is still accelerating in the radio emitting region, if the thermal gradients are assumed to be small. If we ignore thermal gradients completely, then the density coefficients are $p  =2.7$ and 5.5 for $\alpha$ Boo and $\alpha$ Tau, respectively. From mass conservation arguments assuming a steady flow, the power law coefficients for the velocity profiles of each star can be found, ie., $v(r) \propto r^{p-2}$. For $\alpha$ Boo we find $v(r) \propto r^{0.7}$ while for $\alpha$ Tau we find $v(r) \propto r^{3.5}$. This suggests that our long VLA wavelengths may probe a steep acceleration region for $\alpha$ Tau's outflow but for $\alpha$ Boo, may probe a region where the wind is close to its terminal velocity.

@@

The assumption of completely ignoring thermal gradients in a stellar outflow is probably unreliable however. It is likely that some form of Alfv\`{e}n waves are required to lift the material out of the gravitational potential as suggested by \cite{1980ApJ...242..260H}. These waves need damping lengths which are much larger than the density scale height $H_{\star}$ (where $H_{\star} \sim 10 \ H_{\odot}$ for our targets), in order to lift the material out of the gravitational potential but $\lesssim 1 \, R_{\star}$ in order to avoid wind terminal velocities greater than those observed \citep[e.g.][]{1983ApJ...275..808H}. It has also been shown that these waves are expected to produce substantial heating deposition near the base of the wind \citep{1982ApJ...261..279H}. If the long wavelength radio emission from $\alpha$ Tau is indeed emanating from the wind acceleration region, then the dissipation of these Alfv\`{e}n waves may introduce thermal gradients in this region and its velocity profile will not be described by $v(r) \propto r^{3.5}$. Furthermore, diverging flow geometries have been invoked as a more realistic representation of stellar atmospheres \citep{1982ApJ...257..264H,1989A&A...209..327J,2006SSRv..122..181V} and so one could write the continuity equation for a steady flow as 
\begin{equation}
\dot{M} _{\star} = f(r)4\pi r^2 \rho (r) v(r),
\end{equation}
where $f(r)$ is a function describing the divergence from a purely radial flow [i.e., $f(r)=1$], $\dot{M} _{\star}$ is the stellar mass loss rate, and $\rho(r)$ is the gas density at $r$. If the non-radial expansion term can be described by a power-law $f(r) \propto r^s$, where $s>1$ indicates super-radial expansion, then $v(r) \propto r^{3.5-s}$ would describe the velocity profile of $\alpha$ Tau (ignoring thermal gradients and assuming a constant ionization fraction). Further out in the wind, where it has reached its terminal velocity, one would also expect a thermal gradient (but now of opposite sign) due to adiabatic expansion and line cooling. If the long wavelength radio emission emanates from this region of the wind then Equation \ref{eq:eq1} provides us with a direct estimate of the temperature coefficient as we can assume the density coefficient is $p=2$. This may be the case for our long VLA wavelength measurements of $\alpha$ Boo in which case $T_{e}(r) \propto r^{-1.65}$.

@@

1. A constant velocity mass outflow, i.e., $n_{\rm{H}}(r)=C/r^2$ where $C$ is a constant proportional to the ratio of the mass loss rate divided by the terminal velocity. For $\alpha$ Boo, $C = 1.7 \times 10^{32} $ cm$^{-1}$ assuming a wind velocity of 35 km s$^{-1}$.
